---
title: "SRTpipeline - DR-SC Guided Tutorial"
output:
  html_document:
  theme: united
df_print: kable
date: 'Compiled: `r format(Sys.Date(), "%B %d, %Y")`'
---
***

```{r setup, include=FALSE}
all_times <- list()  # store the time for each chunk
knitr::knit_hooks$set(time_it = local({
  now <- NULL
  function(before, options) {
    if (before) {
      now <<- Sys.time()
    } else {
      res <- difftime(Sys.time(), now, units = "secs")
      all_times[[options$label]] <<- res
    }
  }
}))
knitr::opts_chunk$set(
  tidy = TRUE,
  tidy.opts = list(width.cutoff = 95),
  message = FALSE,
  warning = FALSE,
  time_it = TRUE
)
```

## Prepare SRT data for DR-SC model

For this tutorial, we will introduce how to use DR-SC model for joint dimension reduction and spaital clustering based on the R package `DR.SC`.
Here, we will be analyzing  a spatial transcriptomics dataset (SampleID: 151672) for human dorsolateral prefrontal cortex (DLPFC). 
There are 4015  spots and 33538 genes that were sequenced on the 10x Visium platform. Our preprocessed data can be downloaded [here](https://github.com/feiyoung/PRECAST/blob/main/vignettes_data/dlpfc_151672.rda?raw=true), and the raw data can be found [here](https://github.com/LieberInstitute/spatialLIBD).


First, we  loaded the required packages with the command:
```{r init}
library(SRTpipeline)
library(DR.SC)
#library(dplyr)
#library(patchwork)

```

Next, we  download the data to the current working path for the followed analysis by the following command:
```{r eval=FALSE}
githubURL <- "https://github.com/feiyoung/PRECAST/blob/main/vignettes_data/dlpfc_151672.rda?raw=true"
download.file(githubURL,"dlpfc_151672.rda",mode='wb')
```
Then load to R. The data is saved in a Seurat object named `dlpfc_151672`.
```{r  eval = FALSE}
load("dlpfc_151672.rda")
```



<details>
 <summary>**How to read the raw sptial transcriptomics data?**</summary>
In addition, if users' data are in the raw form from spaceranger.
We start by reading in the data. The `Read10X()` function reads in the output of the [spaceranger](https://support.10xgenomics.com/spatial-gene-expression/software/pipelines/latest/what-is-space-ranger) pipeline from 10X, returning a unique molecular identified (UMI) count matrix and a spatial coordinates file. The values in count matrix represent the number of molecules for each feature (i.e. gene; row) that are detected in each spot (column). The spatial information of all spots is represented by the spatial coordinates file.

We next use the count matrix to create a `Seurat` object. The object serves as a container that contains both data (like the count matrix, spatial coordinates) and analysis (like PCA, or clustering results) for a spatial transcriptomics dataset. For a technical discussion of the `Seurat` object structure, check the [GitHub Wiki](https://github.com/satijalab/seurat/wiki). For example, the count matrix is stored in `pbmc[["RNA"]]@counts`.
</details>

<details>
  <summary>**What does data in a count matrix look like?**</summary>

```{r}
# Lets examine a few genes in the first thirty cells

```

The `.` values in the matrix represent 0s (no molecules detected). Since most values in an scRNA-seq matrix are 0,  Seurat uses a sparse-matrix representation whenever possible. This results in significant memory and speed savings for Drop-seq/inDrop/10x data.

````{r}

```
</details>

We check the data by printing it. We can see there are 4015 spots and 33538 genes.
```{r  eval = FALSE}
dlpfc_151672
```
Check the meta data: each row contains the meta information of a spot with row name such as `AAACAAGTATCTCCCA-1`, which consists of the barcode, to identify each spot in RNA sequencing, separateor `-` and a number `1`; note that the columns, `row`, `col`, `imagerow` and `imagecol`, save the spatial information of each spot, in wihch `row` and `col` are the location of a spot on the slide and will be used in DR-SC model, `imagerow` and `imagecol` are the location of a spot on the H&E image; there is the manual annotations `layer_guess_reordered` based on the cytoarchitecture in the original study (Maynard et al., 2021).
```{r  eval = FALSE}
head(dlpfc_151672@meta.data)
```
Check the manual annotations: this slice is annotated as five layers: Layer 3-6 and White matter. 
```{r  eval = FALSE}
table(dlpfc_151672@meta.data$layer_guess_reordered)
```

## Pre-processing workflow

### Create a Seurat object
To show the  pre-processing process of SRT data based on Seurat object, we pick out the count matrix (a sparste matrix object `dgCMatrix`) and meta data (a `data.frame`) in dlpfc_151672, and re-create a Seurat object. 

```{r  eval = FALSE}
meta.data <- dlpfc_151672@meta.data
count_matrix <- dlpfc_151672[['RNA']]@counts
rm(dlpfc_151672)
```

<details>
  <summary>**What does data in a count matrix look like?**</summary>

```{r  eval = FALSE}
# Lets examine a ten genes in the six spots
count_matrix[1:10, 1:4]
```

</details>


First, we change the genes' ensembl name to symbols.
```{r  eval = FALSE}
## Use eg.db database: this method is fast
symbol_name <- transferGeneNames(row.names(count_matrix), now_name = "ensembl", to_name="symbol", species="Human", Method='eg.db')
## Use biomaRt database: this method has less NA but slower than eg.db (about 1 minutes) to finish transfer; and sometimes has network problem when query in the online database. 
#symbol_name <- transferGeneNames(row.names(count_matrix), now_name = "ensembl", to_name="symbol", species="Human", Method='biomaRt')
symbol_name[1:10]
row.names(count_matrix) <- symbol_name
```

Then we create a Seurat object for convenient preprocessing based on Seurat package. 
```{r  eval = FALSE}
library(Seurat)
dlpfc <- CreateSeuratObject(counts=count_matrix, meta.data = meta.data, project='DR_SC')
```
Note: to fit DR-SC model, users require to create a Seurat object with meta.data including spatial coordinates in columns must named “row” (x coordinates) and “col” (y coordinates)!

The SRT data is stored using a default assay, the `RNA`. 
```{r  eval = FALSE}
dlpfc[['RNA']]
```

### QC process
SRTpipeline allows you to easily explore QC metrics and filter cells based on any user-defined criteria. A few QC metrics [commonly used](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4758103/) by the community include
* The number of unique genes detected in each cell. 
    + Low-quality cells or empty droplets will often have very few genes
    + Cell doublets or multiplets may exhibit an aberrantly high gene count
* Similarly, the total number of molecules detected within a cell (correlates strongly with unique genes)


In the example below, we visualize QC metrics, and use these to filter cells.
* filter out the spots with nonzero expression levels for less than 20 genes.
* filter out genes with non-zero expression levels for fewer than 20 spots 
    
```{r qc2,  eval = FALSE, fig.height=7, fig.width=13}
#Visualize QC metrics as a violin plot
VlnPlot(dlpfc, features = c("nFeature_RNA", "nCount_RNA"), ncol = 2)

dlpfc <- filter_gene(dlpfc, min_spots = 20)
dlpfc <- filter_spot(dlpfc, min_feature = 20)
## For this data, all spots passed the QC.
dlpfc
```

### Normalizing the data
After removing unwanted cells and genes from the dataset, the next step is to normalize the data.  Normalized values are stored in `dlpfc[["RNA"]]@data`.
```{r normalize, eval = FALSE}
dlpfc <- NormalizeData(dlpfc)
```

### Feature selection
We next select a subset of genes that exhibit high spot-to-spot variation in the dataset (i.e, they are highly expressed in some spots, and lowly expressed in others). It has been  found that focusing on these genes in downstream analysis helps to highlight biological signal in single-cell datasets [here](https://www.nature.com/articles/nmeth.2645).


#### Choose spatially variable features
We first applied the SPARK-X to choose the top spatially variable genes (SVGs). SPARK-X is embeded in the `DR.SC` R package and is implement in the `FindSVGs()` function. By default, we return 2,000 genes per dataset.  These will be used in downstream analysis, like probabilistic embeddings and spatial clustering. To further speed up the computation, we first use `FindVariableFeatures()` to select 5,000 highly variable genes, then use SPARK-X to choose top SVGs in the  implementation of `FindSVGs()` function. The selected SVGs are saved in `dlpfc@assays$RNA@var.features`.

```{r, eval = FALSE}
dlpfc <- FindSVGs(dlpfc)
dlpfc
```

Users can access the top SVGs using 
```{r, eval = FALSE}
(svg4 <- topSVGs(dlpfc, ntop = 4))
```
Visualize
```{r, eval = FALSE}
dlpfc <- AddCoord2Reduc(dlpfc)
dlpfc <- NormalizeData(dlpfc)


featurePlot(dlpfc, feature = svg4[4], pt_size=2)
```


#### Choose highly variable features
We can also directly use `FindVariableFeatures()` to choose the top highly variable genes (HVGs). We found the performance of using SVGs or HVGs has no significant difference for the downstream analyses, such as dimension reduction and spatial clustering.
```{r, eval=FALSE}
# not run here
#dlpfc <- FindVariableFeatures(dlpfc)
```


## Joint dimension reduction and spatial clustering using DR-SC model
We fit the DR-SC model by using the top 2000 SVGs. Users can also use the top HVGs since we found the performance of using SVGs or HVGs has no significant difference for the downstream analyses, such as dimension reduction and spatial clustering. Here, we specify the number of clusters as 5, while `DR.SC` can also select the number of clusters given a vector for argument $K$.

```{r, eval = FALSE}
### Given K
dlpfc <- DR.SC(dlpfc, K=5, platform = 'Visium', verbose=F)

```

The estimated embeddings are saved in the `dlpfc@reductions$`dr-sc`` and the spatial clusters are saved in the `dlpfc$spatial.drsc.cluster`
```{r, eval = FALSE}
dlpfc@reductions$`dr-sc`
table(dlpfc$spatial.drsc.cluster)
```


## Visualization
Visualize the spatial distribution of cluster labels:
```{r, eval = FALSE}
spatialPlotClusters(dlpfc)
```

```{r, eval = FALSE}
dlpfc <- AddCoord2Reduc(dlpfc)
SpaPlot(dlpfc, item='spatial.drsc.cluster', title_name = NULL)
```


Calculate the 3-dimensional tSNEs based on the extracted features from DR-SC.
```{r, eval = FALSE}
dlpfc <- AddTSNE(dlpfc, reduction = 'dr-sc', assay = 'RNA')
```
We visualize the 3-dimensional tSNEs on the spatial coordinates using RGB colors. We can see there are apparent spatial pattern for the embeddings.
```{r, eval = FALSE}
SpaPlot(dlpfc, item='RGB_tSNE', title_name = "tSNE RGB plot based on DR-SC embeddings")
```
# DEG analysis
After obtain the spatial cluster labels using DR-SC model, we can perform differentially expression analysis. First, we detect the DE genes for cluster 3.
```{r, eval = FALSE}
dat_deg <- FindMarkers(dlpfc, ident.1='cluster3')
subset(dat_deg, avg_log2FC >0.5)
```
We perform differential expression analysis for all clusters by using `FindAllMarkers()` function, then the DE genes' information is saved in a data.frame object `dat_degs`. 
```{r, eval = FALSE}
dat_degs <- FindAllMarkers(dlpfc)
dat_degs
```
We identify the significant DE genes by two criteria: (a) adjustd p-value less than 0.01 and (b) average log fold change greater than 0.4.

```{r, eval = FALSE}
degs_sig <- subset(dat_degs, p_val_adj < 0.01 & avg_log2FC>0.4)
degs_sig
```

In the following, we perform gene set enrichment analysis for the DE genes of each  Domain identified by DR-SC model using R package `gprofiler2`. 
```{r, eval = FALSE}
library(gprofiler2)
termList <- list()
for(k in 1: 5){
  # k <- 1
  cat("k = ", k, '\n')
  dat_degs_sub <- subset(degs_sig,  cluster==paste0("cluster", k))
  
  que1 <- dat_degs_sub$gene
  gostres <- gost(query = que1, organism = "hsapiens", correction_method='fdr')
  termList[[k]] <- gostres
}
termList[[1]]$result
```

 To understand the functions of the identified spatial domains by DR-SC model, we compare the top significant biological process (BP)  pathways in GO database for the DE genes from Domain 1 and 3. Here, we only show to visualize the significant BP pathways and users can explore the other databases such as KEGG and HPA.
 
The enrichment analysis shows that genes specific to Domain 1 were highly enriched for neurofilament bundle assembly,intermediate filament bundle assembly and  cytoskeleton organization, which may indicate the cells in Domain 1 play an important role to achieve these functions. The genes specific to Domain 3 were highly enriched for cytoplasmic translation, ribosome assembly and  regulation of chaperone-mediated autophagy. 
```{r, eval = FALSE}
## Most commonly used databases
source_set <- c("GO:BP","GO:CC", "GO:MF",   'KEGG', "HPA")
cols <- c("steelblue3", "goldenrod", "brown3", "#f98866", "#CE6DBD")
## Here, we show GO:BP
source1 <- "GO:BP"
ss <- which(source_set==source1)
ntop = 5
names(cols) <- source_set
pList_enrich <- list()
for(ii in 1: 5){
  ## ii <- 3
  message("ii=", ii)
  gostres2 <- termList[[ii]]
  dat1 <- subset(gostres2$result, term_size< 500)
  dat1 <- get_top_pathway(dat1, ntop=ntop, source_set = source1)
  dat1 <- dat1[complete.cases(dat1),]
  dat1$nlog10P <- -log10(dat1$p_value)
 
  pList_enrich[[ii]] <-  barPlot_enrich(dat1[order(dat1$nlog10P),], source='source', 'term_name',
                                        'nlog10P', cols=cols[source_set[ss]],base_size=14) + ylab("-log10(p-adj)")+
    xlab("Biological terms") + ggtitle(paste0("Domain", ii))
}
library(patchwork)
wrap_plots(pList_enrich[c(1,3)], nrow=2, ncol=1)

```

We take out the top DE genes for each cluster for visualization. 
```{r, eval = FALSE}

library(dplyr)
n <- 1
dat_degs %>%
  group_by(cluster) %>%
  top_n(n = n, wt = avg_log2FC) -> topGene
topGene
```

```{r, eval = FALSE}

cols_cluster <- c( "#E15759CC", "#59A14FCC", "#FF9DA7CC", "#9C755FCC",
                   "#B07AA1CC")

p1 <- RidgePlot(dlpfc, features=topGene$gene[1:2], cols=cols_cluster, ncol=1)
p1
```


# Trajectory inference



<details>
<summary>**Session Info**</summary>
  ```{r}
sessionInfo()
```
</details>
