---
title: "Getting Started with SRTpipeline"
output:
  html_document:
    theme: united
    df_print: kable
  pdf_document: default
---

```{r fxns, include = FALSE}
library('htmlTable')
make_list <- function(items) {
  paste0("<ul>", sprintf('<li>%s</li>', items), '</ul>', collapse = '')
}
make_href <- function(url, text){
  paste0("<div style='text-align:center; font-size:large; font-weight:bold'><a href=", url ,">", text, "</a></div>")
}
make_href2 <- function(url, text){
  paste0("<a href=", url ,">", text, "</a>")
}
process_entry <- function(dat) {
  if (grepl(pattern = "https://satijalab.org/img/vignette_images", x = dat$image)) {
    img <- paste0('![](', dat$image, '){width=3000px}')
  } else if (grepl(pattern = "assets/", x= dat$image)) {
    img <- paste0('![](', dat$image, '){width=3000px}') 
  } else {
    img <- paste0('![](', '../output/images/', dat$image, '){width=3000px}')
  }
  if (grepl(pattern = "https://satijalab.org/", x = dat$name)) {
    link <- dat$name
  } else {
    link <- paste0(dat$name, ".html")
  }
  go.button <- paste0('<a class="btn btn-primary" href="', link, '" role="button" style="width:100%">GO</a>')
  data.frame(
    title = make_href(url = link, text = dat$title),
    img = img,
    desc = dat$summary,
    btn = go.button
  )
}
process_wrapper_entry <- function(dat) {
  data.frame(
    Package = dat$name,
    Vignette = make_href2(url = dat$link, text = dat$title),
    Reference = make_href2(url = dat$reference, text = dat$citation),
    Source = make_href2(url = dat$source, text = dat$source)
  )
}
make_vignette_card_section <- function(vdat, cat) {
  vignettes <- vdat[[cat]]$vignettes
  dat <- data.frame(title = character(), img = character(), desc = character())
  for (v in 1:length(x = vignettes)) {
    dat <- rbind(dat, process_entry(vignettes[[v]]))
    # dat <- dat[,-2]
    if(nrow(x = dat) == 3 | v == length(x = vignettes)){
      colnames(dat) <- NULL
      dat <- t(dat)
      if (ncol(x = dat) == 2) {
        print(htmlTable(
          dat[-2,], 
          align = '|l|l|', 
          css.cell = "padding-left: .75em; width: 50%", 
          css.class = "two-column-htmltable"
        ))
      } else if (ncol(x = dat) == 1){
        print(htmlTable(
          dat[-2,], 
          align = '|l|', 
          css.cell = "padding-left: .75em; width: 100%", 
          css.class = "one-column-htmltable"
        ))
      } else {
        print(htmlTable(
          dat[-2,], 
          align = '|l|l|l|', 
          css.cell = "padding-left: .75em; width: 30%"
        ))
      }
      dat <- data.frame(title = character(), img = character(), desc = character())
    }
  }
}
```

```{r yaml, include = FALSE}
library(yaml)
vdat <- read_yaml(file = "vignettes.yaml")
```

```{=html}
<style>
.two-column-htmltable {
  width: 66.666%;
}
.one-column-htmltable {
  width: 33.333%;
}

</style>
```

We provide a series of vignettes, tutorials, and analysis walkthroughs to help users get started with SRTpipeline. You can also check out our [Reference page](../reference/index.html) which contains a full list of functions available to users.

# Introductory Vignettes

For new users of SRTpipeline, we suggest starting with a guided walk through of  a spatial transcriptomics dataset (SampleID: 151672) for human dorsolateral prefrontal cortex (DLPFC) from [Github](https://github.com/LieberInstitute/spatialLIBD). This tutorial implements the major components of a standard unsupervised clustering workflow including QC and data filtration, calculation of high-variance genes or spatially variable genes, dimensional reduction, Markov random field-based spatial clustering, and the identification of cluster markers.

We provide additional introductory vignettes for users who are interested in analyzing multiple datasets.

```{r results='asis', echo=FALSE, warning=FALSE, message = FALSE}
make_vignette_card_section(vdat = vdat, cat = 1)
```

# Data Integration

Recently, we have developed [computational methods](https://www.biorxiv.org/content/10.1101/2022.06.26.497672v2) for integrated analysis of SRT datasets generated across different conditions, experimental batches, individuals or datasets with only partially shared cell/domain clusters. As an example, we provide a guided walk through for integrating and comparing hepatocellular carcinoma (HCC) Visium data generated under different  conditions. There are four slides of in-house HCC data generated using the 10x Visium platform, with two slides from tumors (HCC1 and HCC2) and two from tumor-adjacent tissues (HCC3 and HCC4) from an HCC patient. We provide additional vignettes demonstrating how to leverage SC-MEB2 to efficiently integrate large datasets.

```{r results='asis', echo=FALSE, warning=FALSE, message = FALSE}
make_vignette_card_section(vdat = vdat, cat = 2)
```

# Additional New Methods

SRTpipeline also offers additional novel statistical methods for analyzing single-cell data. These include:

* Weighted-nearest neighbor (WNN) analysis: to define cell state based on multiple modalities [[paper](https://doi.org/10.1016/j.cell.2021.04.048)]
* Mixscape: to analyze data from pooled single-cell CRISPR screens [[paper](https://doi.org/10.1038/s41588-021-00778-2)]
* SCTransform: Improved normalization for single-cell RNA-seq data [[paper](https://genomebiology.biomedcentral.com/articles/10.1186/s13059-019-1874-1)]]
* SCTransform, v2 regularization [[paper](https://www.biorxiv.org/content/10.1101/2021.07.07.451498v1.full)]]

```{r results='asis', echo=FALSE, warning=FALSE, message = FALSE}
make_vignette_card_section(vdat = vdat, cat = 3)
```

# Other 

Here we provide a series of short vignettes to demonstrate a number of features that are commonly used in SRTpipeline. We’ve focused the vignettes around questions that we frequently receive from users. Click on a vignette to get started.

```{r results='asis', echo=FALSE, warning=FALSE, message = FALSE}
make_vignette_card_section(vdat = vdat, cat = 4)
```

# SRTpipelineWrappers

In order to facilitate the use of community tools with SRTpipeline, we provide the SRTpipeline Wrappers package, which contains code to run other analysis tools on SRTpipeline objects. For the initial release, we provide wrappers for a few packages in the table below but would encourage other package developers interested in interfacing with SRTpipeline to check out our contributor guide [here](https://github.com/satijalab/SRTpipeline.wrappers/wiki/Submission-Process).

```{r results='asis', echo=FALSE, warning=FALSE, message = FALSE}
library(knitr)
library(kableExtra)
cat <- 5
vignettes <- vdat[[cat]]$vignettes
dat <- data.frame(Package = character(),  Vignette = character(), Reference = character(), Source = character())
for (v in 1:length(x = vignettes)) {
  dat <- rbind(dat, process_wrapper_entry(vignettes[[v]]))
}
dat %>%
  kable(format = "html", escape = FALSE) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```
