---
title: "Installation Instructions for SRTpipeline"
output: html_document
---

To install SRTpipeline, [R](https://www.r-project.org/) version 4.0 or greater is required. We also recommend installing [R Studio](https://www.rstudio.com/).

# Install from CRAN 
 
SRTpipeline is available on [CRAN](https://cran.r-project.org/package=SRTpipeline) for all platforms. To install, run: 

```{r eval = FALSE}
# Enter commands in R (or R studio, if installed)
install.packages('SRTpipeline')
library(SRTpipeline)
```

<!-- If you see the warning message below, enter `y`: -->

<!-- ```{r eval=FALSE} -->
<!-- package which is only available in source form, and may need compilation of C/C++/Fortran: 'SRTpipeline' -->
<!-- Do you want to attempt to install these from sources? -->
<!-- y/n: -->
<!-- ``` -->

# Install the development version of SRTpipeline

Install the development version of SRTpipeline - directly from [GitHub](https://github.com/feiyoung/SRTpipeline/tree/develop).

```{r eval = FALSE}
# Enter commands in R (or R studio, if installed)
# Install the remotes package
install.packages('remotes')
remotes::install_github(repo = 'feiyoung/SRTpipeline', ref = 'develop')
library(SRTpipeline)
```
When users install `SRTpipeline`, the dependent R pacakges will be also installed automatically. 

## Install dependent packages
If dependent packages in CRAN (such as DR.SC) are not normally installed, then use the following command to install the dependent packages.

`install.packages('remotes')`
`remotes::install_github("feiyoung/DR.SC")`

or install from [CRAN](https://cran.r-project.org/)

`install.packages("DR.SC")`


The package can be loaded with the command:
```{r  eval=TRUE}
library("DR.SC")
```



# Installation Instructions

## <i class="fa fa-windows" aria-hidden="true"></i> Installing on Windows

You can install the stable version of `{SRTpipeline}` from CRAN: 

`install.packages("SRTpipeline")`

, or the development version (containing the latest improvements and bug fixes) 
from GitHub:

```{r, eval=FALSE, message=FALSE, warning=FALSE, include=TRUE}
install.packages("remotes")
library(remotes)
install_github("feiyoung/SRTpipeline")
```

## <i class="fa fa-linux" aria-hidden="true"></i> Installing on Linux Systems

To install `{SRTpipeline}` on Linux, you need to be able to install the `{sf}` package, 
which requires several dependencies. See [here](https://github.com/r-spatial/sf#installing)
if you have trouble installing `{sf}`. 

Then, you can install the stable version of SRTpipeline from CRAN:

```{r, eval=FALSE, message=FALSE, warning=FALSE, include=TRUE}
install.packages("SRTpipeline")
```
, or the development version (containing the latest improvements and bug fixes) 
from GitHub:

```{r, eval=FALSE, message=FALSE, warning=FALSE, include=TRUE}
library(devtools)
install_github("feiyoung/SRTpipeline")
```

## <i class="fa fa-apple" aria-hidden="true"></i> Installing on Mac


To install `{SRTpipeline}` on Mac, you need to be able to install the `{sf}` package, 
which requires several dependencies. See [here](https://github.com/r-spatial/sf#installing)
if you have trouble installing `{sf}`. 

Then, you can install the stable version of `{SRTpipeline}` from CRAN:

```{r, eval=FALSE, message=FALSE, warning=FALSE, include=TRUE}
install.packages("SRTpipeline")
```
, or the development version (containing the latest improvements and bug fixes) 
from GitHub:

```{r, eval=FALSE, message=FALSE, warning=FALSE, include=TRUE}
library(devtools)
install_github("feiyoung/SRTpipeline")
```

# Usage

See vigenettes.


# Code of Conduct

Please note that this package is released with a [Contributor Code of Conduct](https://feiyoung.org/code-of-conduct/). 
By contributing to this project, you agree to abide by its terms.
