---
title: "Introduction to SRT integration"
output:
  html_document:
  theme: united
df_print: kable
date: 'Compiled: `r format(Sys.Date(), "%B %d, %Y")`'
---
***

```{r setup, include=FALSE}
all_times <- list()  # store the time for each chunk
knitr::knit_hooks$set(time_it = local({
  now <- NULL
  function(before, options) {
    if (before) {
      now <<- Sys.time()
    } else {
      res <- difftime(Sys.time(), now, units = "secs")
      all_times[[options$label]] <<- res
    }
  }
}))
knitr::opts_chunk$set(
  tidy = TRUE,
  tidy.opts = list(width.cutoff = 95),
  message = FALSE,
  warning = FALSE,
  time_it = TRUE
)
```

# Overview
This tutorial introduce how to use `SRTpipeline` (>=0.1.0) to analyze multiple spatially-resolved transcriptomics (SRT) data.  We emphases how to use `PRECAST` model for spatial embedding, clustering and integration and its followed applications based on `SRTProject` object in the `SRTpipeline` package. 
This tutorial will cover the following tasks, which we believe will be common for many spatial analyses:

* Normalization
* Feature selection
* Dimension reduction
* Spatial clustering
* Data integration
* Batch-corrected gene expression
* DEG analysis
* Spatial trajectory inference
* Visualization

First, we load `SRTpipeline`.
```{r init}
library(SRTpipeline)
set.seed(2023) # set a random seed for reproducibility
```



# Dataset
For this tutorial, we will introduce how to create a `SRTProject` object with multiple SRT samples using SRTpipeline that includes an introduction to common analytical workflows for multiple data batches.
Here, we will be taking spatial transcriptomics dataset for mouse liver as an example.
There are eight tissue slices with 450~700  spots and 8500~9500 genes that were sequenced on the ST platform. Our preprocessed data can be downloaded [here](https://github.com/feiyoung/DR-SC.Analysis/blob/main/data/DLPFC_data/), and the raw data can be found [here](https://github.com/LieberInstitute/spatialLIBD).



Next, we  download the data to the current working path for the followed analysis by the following command:
```{r eval = FALSE}
seuList <- readRDS('seuList_liver8.RDS')                 

```

# Pre-processing workflow
## Prepare SRTProject


Then load to R. 
```{r  eval = FALSE}

n_sample <- length(seuList)
library(Seurat)
## create count matrix list: note each component has a name, i.e., `ID151672`.
cntList <- list()
## create spatial coordinate matrix 
coordList <- list()
## create metadata list
metadataList <- list()

for(r in 1: n_sample){
  # r <- 1
  message('r = ', r)
  seu <- seuList[[r]]
  sp_count <- seu@assays$RNA@counts
  meta.data <- seu@meta.data
  row.names(meta.data) <- colnames(sp_count)
  cntList[[r]] <- sp_count
  metadataList[[r]] <- meta.data
  coordList[[r]] <- cbind(row=seu$row, col=seu$col)
}
names(cntList) <- paste0("Liver", 1:n_sample)
## create meta data for each data batches. Here we only have one data batch.
sampleMetadata <- data.frame(species=rep("Mouse", n_sample),
                             tissues=rep('Liver', n_sample))
row.names(sampleMetadata) <- names(cntList)
## Name of this project
projectName <- "Liver8_PRECAST"
```


## Create a SRTProject object
We start creating SRTProject object. Filter out the genes less than 20 spots with  nonzero read counts and the spots less than 20  genes with nonzero read counts.
```{r  eval = FALSE}
SRTProj <- CreateSRTProject(cntList, coordList, projectName =projectName, metadataList,
    sampleMetadata, min.spots = 20, min.genes = 20)
SRTProj
```


# Normalizing the data
After removing unwanted cells and genes from the dataset, the next step is to normalize the data.  To save RAM memory, normalized values are stored in disk as a `h5file`.
```{r  eval = FALSE}
SRTProj <- normalizeSRT(SRTProj, normalization.method='LogNormalize')
```

# Feature selection
We next select a subset of genes that exhibit high spot-to-spot variation in the dataset (i.e, they are highly expressed in some spots, and lowly expressed in others). It has been  found that focusing on these genes in downstream analysis helps to highlight biological signal in single-cell datasets [here](https://www.nature.com/articles/nmeth.2645).

Then we choose variable features. The default number of variable features is 2,000, and users can change it using argument `nfeatures`. The default type is highly variable genes (HVGs), but users can use spatially variable genes by seting `type='SVGs'`, then `method='SPARK-X'` can be used to choose SVGs.

```{r  eval = FALSE}
SRTProj <- selectVariableFeatures(SRTProj,nfeatures = 2000, type = "HVGs", method='vst')
SRTProj
```

Calculate the adjcence matrix
```{r  eval = FALSE}
## Obtain adjacency matrix
SRTProj <- AddAdj(SRTProj, platform = "ST")

```

# Joint dimension reduction, clustering and integration analysis using PRECAST model 
`PRECAST` model achieves joint dimension reduction, clustering and alignment by integration analysis for multiple samples.
Some SRT clustering methods use Markov random field to model  clusters of spots. These approaches work extremely well and are a standard practice in  SRT data.  `PRECAST` extracts the micro-environment related embeddings and aligned embeddings.
The  micro-environment related  and aligned embeddings are saved in the slot `reductions$microEnv.PRECAST` and `reductions$aligned.PRECAST`, and the spatial clusters are saved in the slot `clusters`.

```{r, eval = FALSE}
SRTProj <- Integrate_PRECAST(SRTProj, K= 6, q = 15)
SRTProj
```


# Visualization
## SRT embeddings
To check the performance of integration, we visuzlize the inferred clusters and data batches on the two-dimensional tSNEs. First, we use `AddTSNE()` function to calculate the two-dimensional tSNEs based on the `aligned.PRECAST` embeddings.
```{r, eval = FALSE}
SRTProj <- AddTSNE(SRTProj, n_comp = 2, reduction='aligned.PRECAST')
```

We visualized the inferred clusters and data batches. The tSNE plot showed the domain clusters were well segregated and data batches were well mixed.
```{r, eval = FALSE}
cols_cluster <- chooseColors(n_colors = 6)
cols_batch <- chooseColors(palettes_name = 'Light 13',n_colors = 8)
p_tsne2_cluster <- EmbedPlot(SRTProj, item='cluster', plotEmbeddings = 'tSNE', cols=cols_cluster, legend.position='right', pt_size=0.2)
p_tsne2_batch <- EmbedPlot(SRTProj, item='batch', plotEmbeddings = 'tSNE', cols=cols_batch, legend.position='right', pt_size=0.2)
drawFigs(list(p_tsne2_cluster, p_tsne2_batch), layout.dim = c(1,2), legend.position = 'right')
```

## Spatial heatmap to show the clustering performance
Except for the embedding plots, SRTpipeline also provides a variaty of visualization functions.
First, we visualize the spatial distribution of cluster labels that shows the layer structure for all data batches.
```{r, eval = FALSE}
## choose colors to function chooseColors
p12 <- EachClusterSpaHeatMap(SRTProj, cols=cols_cluster, legend.position='bottom',base_size=12,
                  pt_size=2, layout.dim=c(2,4), nrow.legend=1)
p12

```

By setting `combine =FALSE`, this function will return a list of ggplot2 objects, thus user can revise each plot. In addtion, we can also plot some of data batches that are interested using the number ID or names of batch.
```{r, eval = FALSE}
pList <- EachClusterSpaHeatMap(SRTProj, cols=cols_cluster, legend.position='bottom',base_size=12,
                  pt_size=0.5, layout.dim=c(2,4), nrow.legend=1, combine=FALSE)
EachClusterSpaHeatMap(SRTProj, batch=1:4, title_name='PRECAST: ',
                      cols=cols_cluster, legend.position='right',base_size=12,
                  pt_size=2, layout.dim=c(2,2))
EachClusterSpaHeatMap(SRTProj, batch= c("Liver1", "Liver2", "Liver4", "Liver6"), title_name='PRECAST: ',
                      cols=cols_cluster, legend.position='right',base_size=12,
                  pt_size=2, layout.dim=c(2,2))
```


## RGB plot to show the embedding performance
Next, we summarized the inferred embeddings for biological eﬀects
between spatial domain types (the slot `reductions`) using three components from either tSNE or UMAP and visualized the resulting tSNE/UMAP components with red/green/blue (RGB) colors in the RGB plot.

The resulting RGB plots from `PRECAST` showed the laminar organization of the human cerebral cortex, and `PRECAST` provided smooth transitions across neighboring spots and spatial domains.
```{r, eval = FALSE}
SRTProj <- AddTSNE(SRTProj, n_comp = 3, reduction='aligned.PRECAST')
p_tsne3<- EachRGBSpaHeatMap(SRTProj, plot_type = "tSNE", pt_size=1 ,title_name="",
                           layout.dim=c(2,4))
p_tsne3
```

To run UMAP in SRTpipeline we use the `AddUMAP()` function. Frist, we evaluate the two-dimensional UMAPs.
```{r, eval = FALSE}
SRTProj <- AddUMAP(SRTProj, n_comp = 2, reduction='aligned.PRECAST')
p_umap2_cluster <- EmbedPlot(SRTProj, item='cluster', plotEmbeddings = 'UMAP', 
                     cols= cols_cluster, legend.position='bottom')
p_umap2_batch <- EmbedPlot(SRTProj, item='batch', plotEmbeddings = 'UMAP', 
                     cols= cols_batch, legend.position='bottom')
drawFigs(list(p_umap2_cluster, p_umap2_batch), layout.dim = c(1,2), 
         legend.position = 'bottom')
```

Then, we evaluate the three-dimensional UMAPs.
```{r, eval = FALSE}
SRTProj <- AddUMAP(SRTProj, n_comp = 3, reduction='aligned.PRECAST')
p_umap3<- EachRGBSpaHeatMap(SRTProj, plot_type = "UMAP", layout.dim=c(2,4),
                        pt_size=1.2 ,title_name="UMAP: ")
p_umap3
```

To save the plot, we can use `write_fig()` function.
```{r, eval = FALSE}
write_fig(p_umap3, filename = 'PRECAST_p_umap3.png',width =14, height =11)
```

Moreover, we can visualize the microenvironment effects on the spatial coordinates.
Then, we evaluate the three-dimensional UMAPs based on `microEnv.PRECAST`.
```{r, eval = FALSE}
# save the previous calculated UMAP3
umap3_cluster <- SRTProj@plotEmbeddings$UMAP3
SRTProj <- AddUMAP(SRTProj, n_comp = 3, reduction='microEnv.PRECAST')
p_umap3_micro <- EachRGBSpaHeatMap(SRTProj, plot_type = "UMAP", layout.dim=c(2,4),
                        pt_size=1.2 ,title_name="mircoEnv: ")
p_umap3_micro
```


## Cell-by-cell heatmap to show the relation of spatial domains
We plotted the heatmap of Pearson’s correlation coefcients of the estimated embeddings among
the detected domains shows the good separation of the estimated embeddings across
domains and the correlations between deeper layers were high, e.g., there were high
correlations between domain 2 and 3, while correlations among the separated layers were low, i.e., domain 1 and 4.

```{r, eval = FALSE}
p_cc <- CCHeatMap(SRTProj, reduction = 'aligned.PRECAST', grp_color=cols_cluster)
p_cc
```

After adding the quantities for data visualization, the `SRTProject` object will have more information in the downstream analyses information.
Now, we print this `SRTProject` object to check it. We observed two components added in the slot `plotEmbeddings` (Embeddings for plotting): `tSNE`, `tSNE3`, `UMAP` and `UMAP3`.
```{r, eval = FALSE}
SRTProj
```




# Combined DEG analysis
After obtain the spatial cluster labels using a clustering model, we can perform differentially expression analysis. 
```{r, eval = FALSE}
speInt <- getIntegratedData(SRTProj, Method='PRECAST', species="Mouse")

```

We perform differential expression analysis for all clusters by using `FindAllMarkers()` function, then the DE genes' information is saved in a data.frame object `dat_degs`.
```{r, eval = FALSE}
dat_degs <- FindAllDEGs(speInt)
dat_degs
```

We identify the significant DE genes by two criteria: (a) adjustd p-value less than 0.01 and (b) average log fold change greater than 0.4.

```{r, eval = FALSE}
degs_sig <- subset(dat_degs, p_val_adj < 0.01 & avg_log2FC>0.15)
degs_sig
```

In the following, we perform gene set enrichment analysis for the DE genes of each  Domain identified by DR-SC model using R package `gprofiler2`.
```{r, eval = FALSE}
library(gprofiler2)
termList <- list()
for(k in 1: 6){
  # k <- 1
  if(sum(degs_sig$cluster==k)> 0){
    cat("k = ", k, '\n')
  dat_degs_sub <- subset(degs_sig,  cluster==k)

  que1 <- dat_degs_sub$gene
  gostres <- gost(query = que1, organism = "hsapiens", correction_method='fdr')
  termList[[k]] <- gostres
  }
  
}
head(termList[[1]]$result)
```

To understand the functions of the identified spatial domains by DR-SC model, we compare the top significant biological process (BP)  pathways in GO database for the DE genes from Domain 1 and 4. Here, we only show to visualize the significant BP pathways and users can explore the other databases such as KEGG and HPA.


```{r, eval = FALSE}
## Most commonly used databases
source_set <- c("GO:BP","GO:CC", "GO:MF",   'KEGG', "HPA")
cols <- c("steelblue3", "goldenrod", "brown3", "#f98866", "#CE6DBD")
## Here, we show GO:BP
source1 <- "GO:BP"
ss <- which(source_set==source1)
ntop = 5
names(cols) <- source_set
pList_enrich <- list()
for(ii in 1: 5){
  ## ii <- 5
  message("ii=", ii)
  gostres2 <- termList[[ii]]
  if(!is.null(gostres2)){
    dat1 <- subset(gostres2$result, term_size< 500)
  dat1 <- get_top_pathway(dat1, ntop=ntop, source_set = source1)
  dat1 <- dat1[complete.cases(dat1),]
  dat1$nlog10P <- -log10(dat1$p_value)

  pList_enrich[[ii]] <-  barPlot_enrich(dat1[order(dat1$nlog10P),], source='source', 'term_name',
                                        'nlog10P', cols=cols[source_set[ss]],base_size=14) + ylab("-log10(p-adj)")+
    xlab("Biological terms") + ggtitle(paste0("Domain", ii))
  }
  
}
drawFigs(pList_enrich[c(1,4)], layout.dim = c(2,1), common.legend = T, align='hv')

```

We take out the top DE genes for each cluster for visualization.
```{r, eval = FALSE}

library(dplyr)
n <- 5
dat_degs %>% as.data.frame %>%
  group_by(cluster) %>%
  top_n(n = n, wt = avg_log2FC) -> topGene
topGene
```

We visualize the DE genes for each cluster group by gene-by-cell heatmap using the `GCHeatMap()` function.
```{r, eval = FALSE}

p1 <- GCHeatMap(speInt, features=topGene$gene, grp_color=cols_cluster ,y_text_size=12)
p1

```


# Trajectory inference
Next, we performed trajectory inference using the aligned embeddings and domain labels estimated by PRECAST model.
```{r, eval = FALSE}
speInt <- AddTrajectory(speInt, reduction = 'aligned.PRECAST')
p1 <- EmbedPlot(speInt, reduction = 'aligned.PRECAST' ,colour_by='PT')
p2 <- EmbedPlot(speInt, reduction = 'tSNE' ,colour_by='PT')
drawFigs(list(p1, p2),layout.dim = c(1,2),common.legend = TRUE, 
         legend.position = 'right')

```

Visualize the inferred pseudotime on the spatial coordinates for each data batch.
```{r, eval = FALSE}
p_spa <- EachEmbedPlot(speInt, reduction = 'Coord' ,colour_by='PT', layout.dim = c(2, 4))
p_spa 
# save(SRTProj,  file=paste0(SRTProj@projectMetadata$outputPath,"/SRTProj.rds"))
# load("F:/Research paper/IntegrateDRcluster/AnalysisCode/SRTpipeline/vignettes/Liver8/SRTProj.rds")

```

# Other downstream analyses


<details>
  <summary>**Session Info**</summary>
```{r}
sessionInfo()
```
</details>
