---
title: "SRT data embedding and clustering"
output:
  html_document:
  theme: united
df_print: kable
date: 'Compiled: `r format(Sys.Date(), "%B %d, %Y")`'
---
***

```{r setup, include=FALSE}
all_times <- list()  # store the time for each chunk
knitr::knit_hooks$set(time_it = local({
  now <- NULL
  function(before, options) {
    if (before) {
      now <<- Sys.time()
    } else {
      res <- difftime(Sys.time(), now, units = "secs")
      all_times[[options$label]] <<- res
    }
  }
}))
knitr::opts_chunk$set(
  tidy = TRUE,
  tidy.opts = list(width.cutoff = 95),
  message = FALSE,
  warning = FALSE,
  time_it = TRUE
)
```

# Overview
This tutorial demonstrates how to use `SRTpipeline` (>=0.1.0) to analyze a spatially-resolved transcriptomics (SRT) data from seqFISH platform. The analytical pipelines are similar to the `SRTpipeline` workflow for Visium SRT analysis. We emphases how to use DR-SC model to achieve spatial embedding and clustering and its followed applications. This tutorial will cover the following tasks, which we believe will be common for many spatial analyses:

* Normalization
* Joint dimension reduction and clustering
* Detecting spatially-variable features
* DEG analysis
* Spatial trajectory inference
* Visualization


First, we load `SRTpipeline`.
```{r init}
library(SRTpipeline)
```


# Dataset
For this tutorial, we will introduce how to create a `SRTProject` object with single SRT sample using SRTpipeline that includes an introduction to common analytical workflows for a single data batch.
Here, we will be taking  a spatial transcriptomics dataset for mouse embryo as an example. 
There are 23194  spots and 351 genes that were sequenced on the seqFISH platform. Our preprocessed data can be downloaded [here](https://github.com/feiyoung/PRECAST/blob/main/vignettes_data/).

We  download the data to the current working path for the followed analysis by the following command:
```{r eval =FALSE}
githubURL <- "https://github.com/feiyoung/PRECAST/blob/main/vignettes_data/Embryo3_seqFISH.rds?raw=true"
download.file(githubURL,"Embryo3_seqFISH.rds",mode='wb')
```
Then load to R. The data is saved in a Seurat object named `Embryo3_seqFISH.rds`.
```{r  eval = FALSE}
load("Embryo3_seqFISH.rds")
```

# Pre-processing workflow

## Create a SRTProject object

First, we show how to create a SRTProject object step by step.

```{r  eval = FALSE}
## create count matrix list: note each component has a name, i.e., `embryo3`.
count_matrix <- seu_em3[['RNA']]@counts
cntList <- list(embryo3=count_matrix)

## create spatial coordinate matrix 
coordList <- list(cbind(seu_em3$x_global, seu_em3$y_global))

## create metadata list
meta.data <- seu_em3@meta.data
metadataList <- list(meta.data)

## create meta data for each data batches. Here we only have one data batch.
sampleMetadata <- data.frame(species="Mouse", tissues=c('Embryo'))
row.names(sampleMetadata) <- names(cntList)

## Name of this project
projectName <- "Embryo3V2"

rm(seu_em3)
```

Next, we start creating SRTProject object. We can print the basic information of this object, including  three parts. The first part have the class of this object, outputPath of data that require to output, h5filePath that save the memory-cusuming data (i.e., count, logcount, ...). The second part is about the datasets basic information, such as how many data batches(sample) and the data names, sample meta data (sampleColData) and meta data for each spot (cellMetaData). The last part is about downstream analyses information that is empty when this object created.
```{r  eval = FALSE}
SRTProj <- CreateSRTProject(cntList, coordList, projectName =projectName, metadataList,
    sampleMetadata)
SRTProj
```   

## Normalizing the data
After creating the SRTProject, the next step is to normalize the data.  To save RAM memory, normalized values are stored in disk as a `h5file`.
```{r  eval = FALSE}
SRTProj <- normalizeSRT(SRTProj)
```
Because seqFISH technology is based on target-genes, we do not select genes but use all the targeted genes by the argument `use_custom_features`. Then the custom features will be regarded as variable features for the followed analyses.
```{r  eval = FALSE}
geneNames <- row.names(SRTProj@geneMetaData)
SRTProj <- selectVariableFeatures(SRTProj, use_custom_features=geneNames)
SRTProj
```

Calculate the adjcence matrix
```{r  eval = FALSE}
## Obtain adjacence matrix
SRTProj <- AddAdj(SRTProj, platform = "Other", radius.upper=12)

```
Visualize
```{r, eval = FALSE}
selectFeatures <- row.names(SRTProj@geneMetaData)[1:2]
EachExprSpaHeatMap(SRTProj, features=selectFeatures, title_name=T, quantVec = c(0.89,0.90))
```








# Spatial embedding and clustering using DR-SC

Some SRT clustering methods use Markov random field to model  clusters of spots, such as SC-MEB and DR-SC. These approaches work extremely well for data with spatial smoothness and are standard practice in SRT data. For this reason, SRTpipeline uses existing state-of-the-art clustering methods from SRT packages for clustering.

We have had the most success using the  Markov random field model implemented by DR-SC. In SRTpipeline, DR-SC joint embedding and clustering is performed using the ` Cluster_DRSC()` function. The resulting embbedings and clusters are saved in `SRTProj@reductions$DR.SC` and `SRTProj@clusters`. The ouput information of `SRTProj` also shows the DR.SC embeddings and inferred cluster labels (Yes).
```{r, eval = FALSE}
SRTProj <- Cluster_DRSC(SRTProj, K= 25, q=15)
SRTProj
```


## Visualization
We visuzlize the inferred cluster labels from `DR.SC` on the two-dimensional tSNEs using `AddTSNE()` and `EmbedPlot()` functions.
```{r, eval = FALSE}
SRTProj <- AddTSNE(SRTProj, n_comp = 2, reduction='DR.SC')
cols_cluster <- c(chooseColors(palettes_name="Blink 23",n_colors = 23), "red4", "green4")
p_tsne2 <- EmbedPlot(SRTProj, item='cluster', plotEmbeddings = 'tSNE', cols=cols_cluster, legend.position='right')
```

To run UMAP in SRTpipeline, we use the `AddUMAP()` function and then vsualize results based on the UMAPs.
```{r, eval = FALSE}
SRTProj <- AddUMAP(SRTProj, n_comp = 2, reduction='DR.SC')
p_umap2 <- EmbedPlot(SRTProj, item='cluster', plotEmbeddings = 'UMAP', cols=cols_cluster, legend.position='right', axis_names=c("UMAP1", "UMAP2"))
p_umap2
```


Next, we merge all plots into one figure.
```{r, eval = FALSE}
p_all <- drawFigs(list(p_tsne2, p_umap2), layout.dim = c(1,2), common.legend = TRUE)
p_all
```

To save the plot, we can use `write_fig()` function.
```{r, eval = FALSE}
write_fig(p_all, filename = 'DRSC_plots.png')
```


Except for the embedding plots, SRTpipeline also provides a variaty of visualization functions.
First, we visualize the spatial distribution of cluster labels that shows the tissue structure.
```{r, eval = FALSE}
EachClusterSpaHeatMap(SRTProj, cols=cols_cluster, legend.position='right',base_size=16, pt_size=0.4)
# remove the border
# EachClusterSpaHeatMap(SRTProj, cols=cols, legend.position='right',base_size=16, border_col='white')
```

We plotted the heatmap of Pearson’s correlation coefcients of the DR.SC embeddings among
the detected spatial domains shows the good separation of the estimated embeddings across
domains and the correlations between deeper layers were high.

```{r, eval = FALSE}
CCHeatMap(SRTProj, reduction = 'DR.SC', grp_color=cols_cluster, ncol.legend=4)
```

After adding the quantities for data visualization, the `SRTProject` object will have more information in the downstream analyses information.
Now, we print this `SRTProject` object to check it. We observed two components added in the slot `plotEmbeddings` (Embeddings for plotting): `tSNE` and  `UMAP`.
```{r, eval = FALSE}
SRTProj
```


# DEG analysis
To do downstream analyses, we require to get the count data from h5file. The function `getGeneSpotData()` access the gene-by-spot count matrix and other data in `SRTProj` and then return a `SpatialExperiment` object,  including two assays: counts, logcounts; rowData: from `SRTProj@geneMetaData`; colData: from `SRTProj@cellMetaData`, `SRTProj@clusters` and sample_id; reducedDim: from `SRTProj@reductions`.
```{r, eval = FALSE}
spe <- getGeneSpotData(SRTProj)
spe
```
We perform differential expression analysis for all clusters by using `FindAllDEGs()` function, then the DE genes' information is saved in a `DataFrame` object `dat_degs`. 
```{r, eval = FALSE}
dat_degs <- FindAllDEGs(spe)
dat_degs
```
We identify the significant DE genes by two criteria: (a) adjustd p-value less than 0.01 and (b) average log fold change greater than 1.

```{r, eval = FALSE}
degs_sig <- subset(dat_degs, p_val_adj < 0.01 & avg_log2FC>1)
degs_sig
```

In the following, we perform gene set enrichment analysis for the DE genes of each  Domain identified by DR-SC model using R package `gprofiler2`. 
```{r, eval = FALSE}
library(gprofiler2)
termList <- list()
for(k in 1: 25){
  # k <- 1
  cat("k = ", k, '\n')
  if(sum(degs_sig$cluster==k)>0){
    dat_degs_sub <- subset(degs_sig,  cluster==k)
  que1 <- dat_degs_sub$gene
  gostres <- gost(query = que1, organism = "hsapiens", correction_method='fdr')
  termList[[k]] <- gostres
  }
  
}
head(termList[[2]]$result)
```

 To understand the functions of the identified spatial domains by DR-SC model, we compare the top significant biological process (BP)  pathways in GO database for the DE genes from Domains 2 and 3. Here, we only show to visualize the significant BP pathways and users can explore the other databases such as KEGG and HPA.
 
```{r, eval = FALSE}
## Most commonly used databases
source_set <- c("GO:BP","GO:CC", "GO:MF",   'KEGG', "HPA")
cols <- c("steelblue3", "goldenrod", "brown3", "#f98866", "#CE6DBD")
## Here, we only show GO:BP
source1 <- "GO:BP"
ss <- which(source_set==source1)
ntop = 5
names(cols) <- source_set
pList_enrich <- list()
for(ii in 1: 25){
  ## ii <- 4
  message("ii=", ii)
  dat <- termList[[ii]]$result
  if(!is.null(dat)){
    dat1 <- subset(dat, term_size< 500)
    dat1 <- get_top_pathway(dat1, ntop=ntop, source_set = source1)
    dat1 <- dat1[complete.cases(dat1),]
    dat1$nlog10P <- -log10(dat1$p_value)
   
    pList_enrich[[ii]] <-  barPlot_enrich(dat1[order(dat1$nlog10P),], source='source', 'term_name',
                                          'nlog10P', cols=cols[source_set[ss]],base_size=14) + ylab("-log10(p-adj)")+
      xlab("Biological terms") + ggtitle(paste0("Domain", ii))
  }
  
}

drawFigs(pList_enrich[c(2:3)], common.legend = T, layout.dim = c(2,1), align = 'hv')

```

We take out the top DE genes for each cluster for visualization. 
```{r, eval = FALSE}
library(dplyr)
n <- 3
dat_degs %>% as.data.frame %>%
  group_by(cluster) %>%
  top_n(n = n, wt = avg_log2FC) -> topGene
topGene
```
We visualize the DE genes for each cluster group by gene-by-cell heatmap using the `GCHeatMap()` function.
```{r, eval = FALSE}
p1 <- EachGCHeatMap(spe, features=topGene$gene, grp_color=cols_cluster ,y_text_size=6, ncol.legend=3)
p1
```


# Trajectory inference
Next, we performed trajectory inference using the PCA embeddings and domain labels estimated by SC-MEB model. The `EmbedPlot()` function can be used to visualize the inferred pseudotime on a specified embedding. If there is only one data batch, the function `EachEmbedPlot()` plays the same role as function `EmbedPlot()`.
```{r, eval = FALSE}
spe <- AddTrajectory(spe, reduction = 'DR.SC', name='PT')
p1 <- EmbedPlot(spe, reduction = 'tSNE' ,colour_by='PT')
p2 <- EmbedPlot(spe, reduction = 'Coord' ,colour_by='PT')
drawFigs(list(p1, p2),layout.dim = c(1,2),common.legend = TRUE, 
         legend.position = 'right')
# EachEmbedPlot(spe, reduction = 'tSNE', colour_by='PT')
# save(SRTProj, spe, file=paste0(SRTProj@projectMetadata$outputPath,"/SRTProj.rds"))
# load("F:/Research paper/IntegrateDRcluster/AnalysisCode/SRTpipeline/vignettes/Embryo3V2/SRTProj.rds")
```



# Other downstream analyses


<details>
<summary>**Session Info**</summary>
```{r}
sessionInfo()
```
